Evaluation of 3-point correlation functions from structural images on CPU and GPU architectures: accounting for anisotropy effects

### 2024-04-25 paper.tex
Fix a typo
Add a citation

To detect the direction of anisotropy we use an $N \times N$ (where $N$ is
dimensionality of the sample) covariance matrix
$\Sigma = \left\{ \sigma_{ij} \right\}$ for coordinates of points in the phase of
interest:
\begin{equation}
  \sigma_{ij} = \sum_{x \in \left\{
    \begin{array}{l}
      \text{Coordinates of points} \\
      \text{in the phase of interset}
    \end{array}
    \right\}} (x_i - \overline{x_i})(x_j - \overline{x_j})
\end{equation}
and decompose $\Sigma$ as $T \Lambda T^{-1}$ where $T$ is a unitary matrix and
$\Lambda$ is diagonal. $T$ can always be made so that it defines a rotation (by
multiplication of one of the columns by $-1$ if needed, so that
$\det T = 1$). Application of this rotation to the sample makes the direction of
anisotropy parallel to one of the axes, so anisotropy can be better represented in
the correlation function \cref{fig:aniso}. In our implementation we use rotation
with nearest-neighbor interpolcation which affects quality of computations as
discussed in \cref{sec:rotqua}.
\begin{figure*}[!htp]
  \centering
  \begin{tabular}{ccc}
    \includegraphics[width=0.3\linewidth, frame]{images/aniso.png} &
    \includegraphics[width=0.3\linewidth, frame]{images/aniso-rot.png} & \\
    \includegraphics[width=0.3\linewidth]{images/aniso-s3.png} &
    \includegraphics[width=0.3\linewidth]{images/aniso-rot-s3.png} &
    \includegraphics[width=0.3\linewidth]{images/aniso-avg-s3.png} \\
  \end{tabular}
  \caption{Behaviour of our algorithm for calculation of the $S_3$ function
    when dealing with highly anisotropic media. The pattern \cref{fig:pattern}
    which we use in our algorithm is most sensitive to anisotropy which is
    directed along one of the axes. \\
    From top down, left to right: an original image of an anisotropic pore; a
    rotated image of the pore, so it is elongated along the axis; $S_3$ function
    for the original pore; $S_3$ function for the rotated pore with a vector
    $(0.92, -0.4)$ being a direction of elongation of the pore; $S_3$ function
    averaged across rotations similar to the algorithm described in
    \cite{berryman1988,SMITH1988176,malmir2018}.}
  \label{fig:aniso}
\end{figure*}

@@

To estimate the error introduced by rotation, consider a two-phase medium
\cref{fig:roterror-noise} whose void phase is defined by an inequality
$f(x, y, z) > U$ where $f$ is a value noise function and $U$ is a threshold. We
produce images of the noise rotated around $Z$ axis by $\theta$ radians by
discretizing an inequation $f(p(x'), p(y'), p(z')) > U$ where
\begin{equation*}
  \begin{aligned}
    x' &= x \cos \theta + y \sin \theta \\
    y' &= y \cos \theta - x \sin \theta \\
    z' &= z
  \end{aligned}
\end{equation*}
and
\begin{equation*}
  \begin{aligned}
    p(x) &= \left\{
    \begin{array}{ll}
      x'     & \quad x' < 1 \\
      2 - x' & \quad \text{otherwise}
    \end{array}
    \right. \\
    x ' &= x \mod 2
  \end{aligned}
\end{equation*}
Therefore we assume the noise to be periodic with the period equal to $2$ and
the interface between phases is continuous. Discretizing this noise for
different values of $\theta \in [0, \pi/4]$ with two different resolutions we
obtain the first dataset which contains the best possible discretization of the
rotated noise at those resolutions. For this study we have produced images with
dimensionality $300 \times 300 \times 300$ and $600 \times 600 \times 600$
voxels (i.e. the second resolution is twice as much as the first).

>>>>

To detect the direction of anisotropy we use an $N \times N$ (where $N$ is
dimensionality of the sample) covariance matrix
$\Sigma = \left\{ \sigma_{ij} \right\}$ for coordinates of points in the phase of
interest:
\begin{equation}
  \sigma_{ij} = \sum_{x \in \left\{
    \begin{array}{l}
      \text{Coordinates of points} \\
      \text{in the phase of interset}
    \end{array}
    \right\}} (x_i - \overline{x_i})(x_j - \overline{x_j})
\end{equation}
and decompose $\Sigma$ as $T \Lambda T^{-1}$ where $T$ is a unitary matrix and
$\Lambda$ is diagonal. $T$ can always be made so that it defines a rotation (by
multiplication of one of the columns by $-1$ if needed, so that
$\det T = 1$). Application of this rotation to the sample makes the direction of
anisotropy parallel to one of the axes, so anisotropy can be better represented in
the correlation function \cref{fig:aniso}. In our implementation we use rotation
with nearest-neighbor interpolation which affects quality of computations as
discussed in \cref{sec:rotqua}.
\begin{figure*}[!htp]
  \centering
  \begin{tabular}{ccc}
    \includegraphics[width=0.3\linewidth, frame]{images/aniso.png} &
    \includegraphics[width=0.3\linewidth, frame]{images/aniso-rot.png} & \\
    \includegraphics[width=0.3\linewidth]{images/aniso-s3.png} &
    \includegraphics[width=0.3\linewidth]{images/aniso-rot-s3.png} &
    \includegraphics[width=0.3\linewidth]{images/aniso-avg-s3.png} \\
  \end{tabular}
  \caption{Behaviour of our algorithm for calculation of the $S_3$ function
    when dealing with highly anisotropic media. The pattern \cref{fig:pattern}
    which we use in our algorithm is most sensitive to anisotropy which is
    directed along one of the axes. \\
    From top down, left to right: an original image of an anisotropic pore; a
    rotated image of the pore, so it is elongated along the axis; $S_3$ function
    for the original pore; $S_3$ function for the rotated pore with a vector
    $(0.92, -0.4)$ being a direction of elongation of the pore; $S_3$ function
    averaged across rotations similar to the algorithm described in
    \cite{berryman1988,SMITH1988176,malmir2018}.}
  \label{fig:aniso}
\end{figure*}

@@

To estimate the error introduced by rotation, consider a two-phase medium
\cref{fig:roterror-noise} whose void phase is defined by an inequality
$f(x, y, z) > U$ where $f$ is a value noise function
\cite[Section IV.B]{samarin2023robust} and $U$ is a threshold. We produce images
of the noise rotated around $Z$ axis by $\theta$ radians by discretizing an
inequation $f(p(x'), p(y'), p(z')) > U$ where
\begin{equation*}
  \begin{aligned}
    x' &= x \cos \theta + y \sin \theta \\
    y' &= y \cos \theta - x \sin \theta \\
    z' &= z
  \end{aligned}
\end{equation*}
and
\begin{equation*}
  \begin{aligned}
    p(x) &= \left\{
    \begin{array}{ll}
      x'     & \quad x' < 1 \\
      2 - x' & \quad \text{otherwise}
    \end{array}
    \right. \\
    x ' &= x \mod 2
  \end{aligned}
\end{equation*}
Therefore we assume the noise to be periodic with the period equal to $2$ and
the interface between phases is continuous. Discretizing this noise for
different values of $\theta \in [0, \pi/4]$ with two different resolutions we
obtain the first dataset which contains the best possible discretization of the
rotated noise at those resolutions. For this study we have produced images with
dimensionality $300 \times 300 \times 300$ and $600 \times 600 \times 600$
voxels (i.e. the second resolution is twice as much as the first).
